---
title: "Homework 1"
author: "Joe Diaz"
date: "9/3/2019"
output: html_document
---


### Question 3.1
**Using the same data set (credit_card_data.txt or credit_card_data-headers.txt) as**
**in Question 2.2, use the ksvm or kknn function to find a good classifier:**


**(a) using cross-validation (do this for the k-nearest-neighbors model; SVM is optional); and**


https://www.rdocumentation.org/packages/kknn/versions/1.3.1/topics/train.kknn

Using the train.kknn function in the kknn library, we can perform leave-one-out crossvalidation easily.
``` {r}
library(kknn)
set.seed(0)
data <- read.table("credit_card_data.txt", stringsAsFactors = FALSE, header = FALSE)

kmax <- 30
model <- train.kknn(V11~.,data,kmax=30,scale=TRUE)
accuracy<- c()
for (k in 1:kmax) {

    model <- cv.kknn(V11~.,data,kcv=10, k=k, scale=TRUE) 
    predicted <- round(model[[1]][,2]) 
    accuracy[k] <- sum(predicted == data$V11)/nrow(data)
}

cat('Acheived best accuracy of: ', round(max(accuracy),2)*100,'%', ' with k=',which.max(accuracy),sep='')

```


**(b) splitting the data into training, validation, and test data sets (pick either KNN or SVM; the other**
**is optional).**
```{r}


idx_train <- sample(nrow(data), size = floor(nrow(data) * 0.6))
train <- data[idx_train,]


rest <- data[-idx_train,]


idx_val <-sample(nrow(rest), size = floor(nrow(rest)*.5))


validation = rest[idx_val,]
test = rest[-idx_val, ] 



```

``` {r}
acc <- c()
for (k in 1:20) {

               # fit k-nearest-neighbor model using training set, validate on test set

     knn_model <- kknn(V11~.,train,validation,k=k,scale=TRUE)

	#  compare models using validation set

     pred <- as.integer(fitted(knn_model)+0.5) # round off to 0 or 1

     acc[k+9] = sum(pred == validation$V11) / nrow(validation)
}

acc[10:29]

## [1] 0.7862595 0.7862595 0.7862595 0.7862595 0.7938931 0.7862595 0.7786260 0.7862595 0.7786260
##[10] 0.7938931 0.7938931 0.7938931 0.7862595 0.7938931 0.7938931 0.8091603 0.8091603 0.8091603
##[19] 0.8091603 0.8091603

# find best-performing KNN model on validation data

cat("Best KNN model is k=",which.max(acc[10:29]),"\n")
cat("Best validation set correctness is ",max(acc[10:29]),"\n")

```

### Question 4.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which a clustering**
**model would be appropriate. List some (up to 5) predictors that you might use.**

### Question 4.2
**The iris data set iris.txt contains 150 data points, each with four predictor variables and one**
**categorical response. The predictors are the width and length of the sepal and petal of flowers and the**
**response is the type of flower. The data is available from the R library datasets and can be accessed with**
**iris once the library is loaded. It is also available at the UCI Machine Learning Repository**
**(https://archive.ics.uci.edu/ml/datasets/Iris ). The response values are only given to see how well a**
**specific method performed and should not be used to build the model.**
**Use the R function kmeans to cluster the points as well as possible. Report the best combination of**
**predictors, your suggested value of k, and how well your best clustering predicts flower type.**
