---
title: "Homework 6"
author: "Joe Diaz"
date: "10/1/2019"
output: pdf_document
---

#### Question 9.1
**Using the same crime data set uscrime.txt as in Question 8.2, apply Principal Component Analysis**
**and then create a regression model using the first few principal components. Specify your new model in**
**terms of the original variables (not the principal components), and compare its quality to that of your**
**solution to Question 8.2. You can use the R function prcomp for PCA. (Note that to first scale the data,**
**you can include scale. = TRUE to scale as part of the PCA function. Don’t forget that, to make a**
**prediction for the new city, you’ll need to unscale the coefficients (i.e., do the scaling calculation in**
**reverse)!)**


Let's read in the table and run pca.
```{r}

df <- read.table("uscrime.txt", stringsAsFactors = FALSE, header = TRUE)
pca <- prcomp(df[,1:15], scale. = TRUE)

components <- pca$x[,1:4] # store key columns as a dataframe
summary(pca)

```

build linear model on pca
```{r}
crime_components <- cbind(components, df[,16])
model <- lm(V5~., data = as.data.frame(crime_components))
#summary(model)

intercept <- model$coefficients[1]
coeff <- model$coefficients[2:5]

```

reverse pca

```{r}
alphas <- pca$rotation[,1:4] %*% coeff


origAlpha <- alphas/sapply(df[,1:15],sd)

m <- sapply(df[,1:15],mean)
std <- sapply(df[,1:15],sd)

intercept - sum(coeff*m/std)

origIntercept<- sum(coeff*sapply(df[,1:15],mean)/sapply(df[,1:15],sd))

```
generate predictions on original alphas and betas
```{r}
preds <- as.matrix(df[,1:15]) %*% origAlpha + origIntercept



sse = sum((preds - df[,16])^2)
tot_ss = sum((df[,16] - mean(df[,16]))^2)



r2 <- 1 - sse/tot_ss

r2
```
achieved r2 of .31