---
title: "Homework 1"
author: "Joe Diaz"
date: "8/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**Question 2.1**

Describe a situation or problem from your job, everyday life, current events, etc., for which a
classification model would be appropriate. List some (up to 5) predictors that you might use.


**Question 2.2**

The files credit_card_data.txt (without headers) and credit_card_data-headers.txt
(with headers) contain a dataset with 654 data points, 6 continuous and 4 binary predictor variables. It
has anonymized credit card applications with a binary response variable (last column) indicating if the
application was positive or negative. The dataset is the “Credit Approval Data Set” from the UCI Machine
Learning Repository (https://archive.ics.uci.edu/ml/datasets/Credit+Approval) without the categorical
variables and without data points that have missing values.


1. Using the support vector machine function ksvm contained in the R package kernlab, find a
good classifier for this data. Show the equation of your classifier, and how well it classifies the
data points in the full data set. (Don’t worry about test/validation data yet; we’ll cover that
topic soon.)

```{r cars}
#install.packages('kernlab')
library('kernlab')


data <- as.matrix(read.table('credit_card_data.txt', sep = '\t'))


# call ksvm. Vanilladot is a simple linear kernel.
model <- ksvm(data[,1:10],data[,11],type="C-svc",kernel="vanilladot",C=100,scaled=TRUE)

# calculate a1…am
a <- colSums(model@xmatrix[[1]] * model@coef[[1]])
a
# calculate a0
a0 <- -model@b
a0
# see what the model predicts
pred <- predict(model,data[,1:10])
pred
# see what fraction of the model’s predictions match the
# actual classification
sum(pred == data[,11]) / nrow(data)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
