---
title: "Homework 5"
author: "Joe Diaz"
date: "9/25/2019"
output: pdf_document
---

#### Question 8.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which a linear**
**regression model would be appropriate. List some (up to 5) predictors that you might use.**

A nice use case for a linear regression would be predicting the value of a used car to see if you are getting a good deal or not.

Some features I would use are:

- Make
- Model
- Age
- color
- Location of vendor


#### Question 8.2
**Using crime data from http://www.statsci.org/data/general/uscrime.txt (file uscrime.txt,**
**description at http://www.statsci.org/data/general/uscrime.html ), use regression (a useful R function is**
**lm or glm) to predict the observed crime rate in a city with the following data:**
```{r}

#df <- read.table("uscrime.txt", stringsAsFactors = FALSE, header = TRUE)
df<-read.delim("http://www.statsci.org/data/general/uscrime.txt")


model <- glm(Crime ~ . , data=df, family="gaussian")
summary(model)

```


## Model Validation
I referenced this to perform cross-validation with glm https://stat.ethz.ch/R-manual/R-patched/library/boot/html/cv.glm.html
```{r}
library(boot)
# sum of squared differences
sq_diff <- sum((df$Crime - mean(df$Crime))^2)

cross_val_model <- cv.glm(df,model,K=7)

# R squared
1 - cross_val_model$delta[1]*nrow(df)/sq_diff
```

R-Squared is fairly low when using all the available features!